Promotors' reports on the first draft of the thesis document.

### 2020-08-21 ChapterIntroduction/ChapterIntroduction.tex
Second draft before public defense

% Context:Metrology One of the objectives in metrology is the estimation of the true value of physical quantities.
A measurement is a dynamic process, and a sensor is a dynamic system.
The physical quantities of interest are the inputs of the system, and the outputs are the electrical signals collected from the sensor.
The inputs interact with the sensor, and there are energy transfers between them that modify the sensor state.
The sensor output response depends on the applied inputs and on the sensor initial conditions.

@@

% Context:Data-driven step input estimation method
An example of a data-driven method devised for a DSP is the direct step input level estimation from the sensor step response \citep{Markovsky15cep}.
This method formulates a Hankel structured errors-in-variables (EIV) problem with correlation.
The regression matrix has a block-Hankel structure.
The correlation exists because the transient response, perturbed by measurement noise, constructs both the regression matrix and the regressor.
The measurement noise is assumed to have zero mean and finite variance.
The method is implemented in real-time using a recursive least-squares (RLS) solution of the structured EIV problem.
% Avoiding the model identification from input-output data and directly estimating the input from the transient response reduces the input estimation time and makes data-driven input estimation methods suitable for real-time

% Context:Data-driven methodology
% The philosophy behind the estimation method is avoiding the explicit identification of a system model from input-output data and estimating directly the input from the transient response.
The main advantage of the data-driven step input estimation method is that it does not identify the sensor model, but instead, it directly estimates the input. 
The direct estimation differs from the standard two-stage methodology, that first identifies the sensor model and later estimates the input.
In this approach, the output-error (OE) problem is converted into an EIV problem that is harder to solve, but the RLS solution is easy to compute.
The range of application of the data-driven input estimation method is extensive because it is independent of the sensor model.
The main disadvantage of the data-driven input estimation method is that its stochastic properties are not straightforwardly evident. 
It is more complex to find the stochastic properties of EIV problems when they have structure and correlation.
% Moreover, the online uncertainty assessment may not be feasible and we have to rely on confidence bounds.

% Problem under study: Dynamic measurements
% DM:Uncertainty assessment of data-driven input estimation methods
% DM:Experimental validation of the data-driven step input estimation method
The estimation uncertainty assessment is needed to validate the estimation methods for metrology applications.
The uncertainty of the data-driven step input estimation method \citep{Markovsky15ieee} is unknown.
The estimation bias and variance define the estimation uncertainty \citep{Pintelon12Book}, and they can be obtained by conducting an elementwise statistical analysis. 
The validation of the step input estimation method requires also to demonstrate its effectiveness on real-life measurements.
Temperature and mass sensors are suitable for real-life measurement experiments.
One challenge of using real-life sensor outputs is that the noise may not fulfill the whiteness assumed in the estimation problem formulation and in the statistical analysis.
The experimental and simulation results permit to evaluate the step input estimation method performance.

@@

The scientific literature contains metrology studies that deal with the dynamic process effects of measurements.
This subsection presents a list of relevant works published in relation with the thesis topic. 
These works study theoretical aspects like the estimation of inherent dynamical errors \citep{Hessling06}, the dynamic correction in time domain \citep{Hessling08a}; and experimental aspects like the increment of the amount of sensors in production lines for quick decision making in autonomous control \citep{Esward09}.

To implement input estimation methods in the practice, the compensation filters based on deconvolution \citep{Eichstadt10} are the preferred methodology to develop 
analog filters \citep{Jafaripanah05}, 
adaptive digital filters \citep{Shu93}, 
lattice adaptive filters \citep{Hernandez06}, 
compensators for simultaneous responses of different sensors \citep{Boschetti13}, 
regularized deconvolution compensators \citep{Dienstfrey14}, 
and a multiple choice set of filters \citep{Huang16}. 
All these methods have in common the use of the measurement system model parameters to build the compensator.

% Digital signal processing estimation methods
A list of measurement methods based on digital signal processors (DSP) include 
a study of the data-driven dynamic error correction and its impact on the measured temperature \citep{Saggin01},
a modulation quality measurement of microwave access systems \citep{Angrisani10}, 
a real-time rotational speed estimation using correlation \citep{Wang14},
a development of a biology-inspired electronic nose \citep{Jing16},
and impedance measurements for material damage estimation using cross-correlation \citep{deCastro19}.
The data-driven step input estimation method \citep{Markovsky15cep} is added to this list, highlighting the fact that it can be implemented for the measurement of different physical magnitudes because it is model-independent.

% Uncertainty assessment 
In metrology, a measurement is an estimation, represented as a random variable.
The measurement noise always exists, and therefore, the input estimation is commonly expressed with its two first statistical moments \citep{Ferrero06}.
The guideline \citep{GUM08}, accepted by the metrology community, recommends the standardization of the uncertainty assessment.
The typical measurement uncertainty analysis are reviewed in \citep{daSilva12}.
The Monte Carlo method is a uncertainty evaluation tool \citep{Cox06},
that supports the use of simulation techniques for quantifying measurement uncertainties \citep{Esward16}.
Another example of the Monte Carlo method application is a dynamic measurement uncertainty evaluation of clinical thermometers \citep{Ogorevc16}.

Nevertheless, researchers still pinpoint the need to study more the uncertainty assessment methods \citep{Esward09, Hessling10}.
Some recommend to consider the uncertainties of all the measurement chain components \citep{Diniz17}, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.  
Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated for
a discrete-time infinite-response filter in \citep{Link09},
a discrete-time finite-response filter in \citep{Elster07, Elster08}, and
the Kalman filter in \citep{Eichstadt16b}.
All these works propagate the uncertainty through the filter, but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions \citep{Hessling11}.

@@

This thesis describes research work initiated after the data-driven step input estimation method \citep{Markovsky15cep}.
This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
The first research work conducted was a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
After obtaining these statistical moments, the estimation uncertainty was assessed, and thus, the effectiveness of the method was appraised.
The second research work was a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
The experiments were realizations of the step input excitation using temperature and mass sensors.
The step input estimation method showed robustness when the measurement noise is not Gaussian and white, as it was assumed in the theoretical analyses.
The third research work conducted was the estimation of affine input parameters following the data-driven approach for real-time applications. 
The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum-likelihood method.

@@

The data-driven step input estimation method formulates a structured errors-in-variables (EIV) problem.
In linear estimation EIV problems, the measurement noise perturbs the regression matrix and the regressor \citep{VanHuffel91Book}, \citep{Markovsky07overview}.
The regression matrix, in structured EIV problems, has a structure that depends on the problem formulation.
The Hankel and Toeplitz matrices appear in problems of metrology \citep{Markovsky15cep}, system identification \citep{Soderstrom07}, image restoration \citep{Feiz17}, nuclear magnetic resonance spectroscopy \citep{Cai16}, direction-of-arrival estimation \citep{Pan18}, and time-of-arrival estimation \citep{Jia18}.

The data-driven step input estimation method directly estimates the input from the sensor transient response. 
The perturbations in the EIV problem of interest come from the sensor output.
This is the only observed signal \citep{Markovsky15cep}, and it constructs both the regression matrix and the regressor.
The structure in the regression matrix is block-Hankel.
The data-driven direct estimation methodology  estimates the input faster than the classical two-stage approach, which first identifies a sensor model and later estimates the input using the sensor model \citep{Azam15, Niedzwiecki16a}.

Instead of using a total least-squares (TLS) solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
The TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed \citep{Markovsky07overview}. 
However, regarding structured EIV problems, the TLS estimator does not give general results since each structure requires a specific treatment \citep{VanHuffel07TLSeditorial}, and the ML estimator leads to non-convex optimization problems where finding the global optimum is not guaranteed \citep{Rhode14recursive}.
Moreover, the computational complexity of TLS and ML inhibits their real-time implementation.
The least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form (RLS) suitable for real-time implementation.

The previously published works that propose LS estimators to solve structured EIV problems do not study the required statistical moments to know the uncertainty of the data-driven step input estimation method.
In the literature we find 
the design of a fast algorithm for matrices with small displacement rank \citep{Mastronardi07fast}, 
the study of the LS estimator consistency \citep{Palanthandalam10parameter},
the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models \citep{Kiviet12high}, \citep{Kiviet14improved}, and
a discussion of the causes of bias and inconsistency in homogeneous estimators \citep{Yeredor04homogeneous}.
The literature does not address the uncertainty of LS estimators for structured EIV problems.

The uncertainty of an estimation method is expressed using the bias and covariance of the estimate \citep{Pintelon12Book}.
To know the uncertainty of the data-driven step input estimation method, I quantified the bias and covariance of the LS solution of EIV problems, for unstructured and structured cases. 
The bias and covariance quantification extend the perturbation analysis that investigated in \citep{Stewart90SPT} and in \citep{Vaccaro94}.
The conducted analysis provides insight into the impact that the structure and the correlation have on the LS estimation uncertainty.
The study presented in this thesis illustrates a methodology to conduct statistical analysis for any structured EIV problem.

I derived expressions that quantify the bias and covariance by obtaining the mathematical expectation of the LS estimate approximated by the second-order Taylor series expansion.
Using Monte Carlo simulations, I validated the accuracy of the bias and covariance expressions.
These expressions estimate the bias and covariance that the data-driven step input estimation method will exhibit for a given sample size and perturbation level.
I compared the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions under which the data-driven step input estimation method is appropriate for practical applications \citep{QuintanaCSDA}.

@@

The validation of the data-driven step input estimation method in a practical application was necessary to demonstrate the usefulness of the method.
The method performance was illustrated by simulations and temperature experiments on a digital signal processor (DSP) of low cost \citep{Markovsky15cep}.
The method estimates the unknown level of step inputs by processing the sensor step response, and 
avoiding the sensor modeling stage.
The formulation of the estimation method is a correlated errors-in-variables (EIV) problem with block-Hankel structure.

Other methods for input estimation mainly compensate the sensor transient response, for example, by 
recursive estimation of the compensator parameters \citep{Shu93}, 
finite impulse response (FIR) filtering \citep{Elster07}, \citep{Niedzwiecki16b} filters and 
infinite impulse response (IIR) filtering \citep{Pintelon90}, \citep{Elster08}.
The uncertainty propagation for these model-based compensators is based on the transfer function or state-space representations of the LTI sensor and filter systems \citep{Link09}, \citep{Hale09}.
Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citep{Pietrzak14} for mass and in \citep{Ogorevc16} for temperature sensors.

To validate the data-driven step input estimation method I built a weighing system setup based on a load cell sensor.
The weighing setup was constructed to ensure repeatability and reproducibility, along with different experimental realizations.
Load cell sensors are versatile devices that are found in 
heart and breathing physiological signal monitoring \citep{Lee16},
clinical analysis of sleep quality \citep{Zahradka18},
automobile safety studies \citep{Ballo16},
wind turbine design \citep{Rossander15}, 
civil engineering \citep{Olmi16}, and 
sport bicycle design \citep{Casas16}, to name a few.

\begin{comment}
% Load cell sensors 
Weighing has been basic for the development of scientific and trade activites. 
The load cell is now a standard transducer for weight determination and also for the improvement of measurement techniques, such as the geometric approach to processing of load cell responses \citep{Kesilmis16}, the design of new conveyor machinery \citep{Yamani18}, and electronic truck scales \citep{Guo18}.

In safety studies, a six axis load cell is devised to quantify accelerations and impact forces exerted on a dummy \citep{Ballo16}.
In alternative energy developments the load cells are useful to measure the forces on the arms of a vertical axis wind turbine \citep{Rossander15}.
An academic study of the load that a structure withstands is conducted with strain gage load cells that confirms the numerical results and facilitates the design of complex shaped structures \citep{Olmi16}.
In sports, the performance of new instrumented crank mechanisms is fostered by the utilization of load cells in the characterization, analysis and validation design stages \citep{Casas16}.

The versatility of the load cells permits the physiological signal monitoring of the the heart and breathing rates \citep{Lee16}, clinical analysis of sleep quality \citep{Zahradka18} and the classification of the movement intensity of people while they are sleeping \citep{Alaziz17}. 
All of these experimental studies have load cells installed on bed setups.

>>>>

% Context:Metrology One of the objectives in metrology is the estimation of the true value of physical quantities.
A measurement is a dynamic process, and a sensor is a dynamic system.
The physical quantities of interest are the inputs of the system, and the outputs are the electrical signals collected from the sensor.
The inputs interact with the sensor, and there are energy transfers between them that modify the sensor state.
The sensor output response depends on the applied inputs and on the sensor's initial conditions.

\color{blue}

Take for example the case of a thermometer.
To measure temperature, the thermometer is put in contact with the substance or object of interest.
According to the Newton's law of cooling, the temperature $y$ of the thermometer changes proportionally to its difference with the temperature $u$ of the matter under study:
\begin{equation}  \dfrac{dy}{dt} = k_{\mathrm{T}} \left( u - y \right) \label{eqn:NewtonCooling} \end{equation}
where the proportionality constant $k_{\mathrm{T}}$ is a parameter of the thermometer's first order model.
The input $u$ is unknown and its estimation is the objective of the measurement.
It is customary to take the reading from the thermometer once the temperature stops changing.

Another popular example of a measurement is found in weighing. 
In most of modern scales, the object of interest is placed on top of a weighing sensor.
A simple model of the weighing sensor is a second order mass-spring-damper model shown in Figure \ref{fig:simple_msd_system}.
In this model, the unknown mass $u(t)$, that can be time-dependent, causes a displacement $y$ of the scale and the dynamics of the sensor are described by the differential equation:
\begin{equation} \dfrac{d}{dt} \left( \left(u(t)+m\right) \ \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left(u(t)+m\right)  g \end{equation}
where $m$ is the own mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.

For illustration purposes, Figures \ref{fig:thermometer} and \ref{fig:scale} show the responses of the thermometer and the weighing sensors, respectively, after the application of a constant input $u$, which can be represented by a step input.

\begin{figure}[htb!]
\centering

\begin{tikzpicture}[every node/.style={draw,outer sep=0pt,thick}]
\tikzstyle{spring}=[thick,decorate,decoration={zigzag,pre length=0.3cm,post length=0.3cm,segment length=6}]
\tikzstyle{damper}=[thick,decoration={markings,  
  mark connection node=dmp,
  mark=at position 0.5 with 
  {
    \node (dmp) [thick,inner sep=0pt,transform shape,rotate=-90,minimum width=15pt,minimum height=3pt,draw=none] {};
    \draw [thick] ($(dmp.north east)+(2pt,0)$) -- (dmp.south east) -- (dmp.south west) -- ($(dmp.north west)+(2pt,0)$);
    \draw [thick] ($(dmp.north)+(0,-5pt)$) -- ($(dmp.north)+(0,5pt)$);
  }
}, decorate]
\tikzstyle{ground}=[fill,pattern=north east lines,draw=none,minimum width=0.63cm,minimum height=0.3cm]

\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};

\node (ground1) at (M.south) [ground,yshift=-1.5cm,xshift=-0.625cm,anchor=north] {};
\draw (ground1.north west) -- (ground1.north east);
\draw [spring] (ground1.north) -- ($(M.south east)!(ground1.north)!(M.south west)$);

\node (groundc) at (M.south) [ground,yshift=-1.5cm,anchor=north] {}; 
\draw (groundc.north west) -- (groundc.north east);

\node (ground2) at (M.south) [ground,yshift=-1.5cm,xshift=0.625cm,anchor=north] {};
\draw (ground2.north west) -- (ground2.north east);
\draw [damper] (ground2.north) -- ($(M.south east)!(ground2.north)!(M.south west)$);

\node[draw=none,fill=none] at (-0.9cm,-1cm) {$k_{\mathrm{s}}$};
\node[draw=none,fill=none] at (0.15cm,-1cm) {$k_{\mathrm{d}}$};
\node[draw=none,fill=none] at (2.0cm,1.0cm) {$y$};
\draw [-latex,thick]  ++(2.2cm,-1cm) -- +(0cm,2.25cm);

\draw [-latex,thick] (M.east) ++(0,0) -- +(1cm,0);
\draw [line width=0.25mm] (2.2cm,-1cm) -- (2.2cm,1cm);
\draw [line width=0.25mm] (2.1cm,-1cm) -- (2.3cm,-1cm);
\draw [line width=0.25mm] (2.1cm,1cm) -- (2.3cm,1cm);
\draw [line width=0.25mm] (2.1cm,-0.5cm) -- (2.3cm,-0.5cm);
\draw [line width=0.25mm] (2.1cm,0.5cm) -- (2.3cm,0.5cm);
\draw [line width=0.25mm] (2.15cm,-0.25cm) -- (2.25cm,-0.25cm);
\draw [line width=0.25mm] (2.15cm,0.25cm) -- (2.25cm,0.25cm);
\draw [line width=0.25mm] (2.15cm,-0.75cm) -- (2.25cm,-0.75cm);
\draw [line width=0.25mm] (2.15cm,0.75cm) -- (2.25cm,0.75cm);
\draw [line width=0.25mm] (2.1cm,0cm) -- (2.3cm,0cm);

\end{tikzpicture}

\caption{\label{fig:simple_msd_system} \color{blue} A second order mass-spring-damper model represents the weighing sensor. The application of the mass $u$ causes a change in the position $y$ of the scale. \color{black} } 
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_2.pdf} 
\caption{ \label{fig:thermometer} 
\color{blue} This simulation shows an example of the temperature change experimented by a thermometer, with $k_{\mathrm{T}}=0.5 s^{-1}$, caused by a step input of 67 $^{\circ} \mathrm{C}$  when the thermometer was initially at 23 $^{\circ} \mathrm{C}$. In this example, we should wait 10 s to have an estimation of the input with a relative error lower than 0.5 \%. \color{black} }
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_3.pdf} 
\caption{ \label{fig:scale} 
\color{blue} This simulation illustrates the displacement observed when a constant input $u=1 \mathrm{kg}$ is applied to a weighing sensor in equilibrium. The sensor parameters are $k_{\mathrm{d}} = 5 \ \mathrm{N \ s/m}$, $m = 5 \ \mathrm{g}$, and $k_{\mathrm{s}} = 10250 \ \mathrm{N/m}$. In this conditions, the user has to wait more that 2 s to observe a stabilization on the scale and to estimate the mass from reading the final value of the displacement with a relative error smaller than 0.72 \%. \color{black}  }
\end{figure}

In \citet{Eichstadt12} an illustration on a force measurement sensor can be found. 
The sensor is modeled as a cascade of two damped mass-spring systems. 
The transfer function of the sensor is $H(s) = H_1(s) \ H_2(s)$, where each of the two transfer functions are described by
\begin{equation} H_i(s) = \dfrac{\left( 2 \pi f_{0,i}\right)^2}{s^2  + 2 \delta_i \left( 2 \pi f_{0,i}\right) s  +  \left( 2 \pi f_{0,i}\right)^2} 
\end{equation}  
where $\delta_1 = 0.1$, $\delta_2 = 0.2$, $f_{0,1} = 1.9$ Hz, and $f_{0,2} = 7$ Hz are the system parameters.
The response of this sensor to a unit step input, shown in Figure \ref{fig:force}, takes at least 5 s to in which the transient response is predominant.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_4.pdf} 
\caption{ \label{fig:force} 
\color{blue} The simulation of a force sensor based on a double mass-spring system in cascade also shows that the stabilization after a unit step input is not instantaneous. In the plot we infer that the user has to wait 5 s to estimate the force from the sensor response that is 0.25 \% far from its value in equilibrium. \color{black}  }
\end{figure}

The handbook of sensors written by \citet{Fraden16Book} describes a pressure sensor that detects sudden changes in the pressure inside a room.
These changes can be caused by the opening of a door or by the movement of a person.
The principle of operation of detecting minute pressure gradients is the flowmeter that compares the pressure at both outlets of a tube.   
An experimental identification of such a flowmeter is described in \citet{Wiklund02}, where a second order model is proposed to explain the dynamics with the transfer function
\begin{equation} H(s) = \dfrac{ e^{-\tau_{delay}s}}{\left( 1 +  \tau_1 s \right) \left( 1 +  \tau_2 s \right)} 
\end{equation}  
where $\tau_{delay} = 0.07$ s is the dead time, and $\tau_{1} = 0.01$ s and $\tau_{2} = 1.6$ s are two time constants.
The step response simulation of this dynamic model results in the sensor step response shown in Figure \ref{fig:pressure}.  
We have a response to a unit step function that is mostly and increasing exponential that requires 7 s approximately to reach a value that is 98.7 \% ot its maximum value.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterIntroduction/fig/Fig_5.pdf} 
\caption{ \label{fig:pressure} 
\color{blue} The response of a flowmeter to a sudden change in its pressure gradient acrros the extremes of a tube is an overdamped transient response. Ideally we would expect the the transient response follows closely the shape of the unit step input applied, but instead an exponential response is observed, that last for more than 7 s.  \color{black}  }
\end{figure}

\color{black}

@@

% Context:Data-driven step input estimation method
An example of a data-driven method devised for a DSP is the direct step input level estimation from the sensor step response introduced in \citet{Markovsky15cep}.
This method formulates a Hankel structured errors-in-variables (EIV) problem with correlation.
The regression matrix has a block-Hankel structure.
The correlation exists because the transient response, perturbed by measurement noise, constructs both the regression matrix and the regressor.
The measurement noise is assumed to have zero mean and finite variance.
The method is implemented in real-time using a recursive least-squares (RLS) solution of the structured EIV problem.
% Avoiding the model identification from input-output data and directly estimating the input from the transient response reduces the input estimation time and makes data-driven input estimation methods suitable for real-time

% Context:Data-driven methodology
% The philosophy behind the estimation method is avoiding the explicit identification of a system model from input-output data and estimating directly the input from the transient response.
The main advantage of the data-driven step input estimation method is that it does not identify the sensor model, but instead, it directly estimates the input. 
The direct estimation differs from the standard two-stage methodology that first identifies the sensor model and later estimates the input.
In this approach, the output-error (OE) problem is converted into an EIV problem that is harder to solve, but the RLS solution is easy to compute.
The range of application of the data-driven input estimation method is extensive because it is independent of the sensor model.
The main disadvantage of the data-driven input estimation method is that its stochastic properties are not straightforwardly evident. 
It is more complex to find the stochastic properties of EIV problems when they have structure and correlation.
% Moreover, the online uncertainty assessment may not be feasible and we have to rely on confidence bounds.

% Problem under study: Dynamic measurements
% DM:Uncertainty assessment of data-driven input estimation methods
% DM:Experimental validation of the data-driven step input estimation method
The estimation uncertainty assessment is needed to validate the estimation methods for metrology applications.
The uncertainty of the data-driven step input estimation method described in \citet{Markovsky15ieee} is unknown.
The estimation bias and variance define the estimation uncertainty, as it is explained in \citet{Pintelon12Book}, and they can be obtained by conducting an elementwise statistical analysis. 
The validation of the step input estimation method requires also to demonstrate its effectiveness on real-life measurements.
Temperature and mass sensors are suitable for real-life measurement experiments.
One challenge of using real-life sensor outputs is that the noise may not fulfill the whiteness assumed in the estimation problem formulation and in the statistical analysis.
The experimental and simulation results permit to evaluate the step input estimation method performance.

@@

The scientific literature contains metrology studies that deal with the dynamic process effects of measurements.
This subsection presents a list of relevant works published in relation with the thesis topic. 
These works study theoretical aspects like the estimation of inherent dynamical errors presented in \citet{Hessling06}, the dynamic correction in time domain proposed in \citet{Hessling08a}; and experimental aspects like the increment of the amount of sensors in production lines for quick decision making in autonomous control described in \citet{Esward09}.

To implement input estimation methods in the practice, the compensation filters based on deconvolution explained by \citet{Eichstadt10} are the preferred methodology to develop, for example, 
analog filters in \citet{Jafaripanah05}, 
adaptive digital filters in \citet{Shu93}, 
lattice adaptive filters in \citet{Hernandez06}, 
compensators for simultaneous responses of different sensors in \citet{Boschetti13}, 
regularized deconvolution compensators in \citet{Dienstfrey14}, 
and a multiple choice set of filters in \citet{Huang16}. 
All these methods have in common the use of the measurement system model parameters to build the compensator.

% Digital signal processing estimation methods
A list of measurement methods based on digital signal processors (DSP) include 
a data-driven dynamic error correction and its impact on the measured temperature studied in \citet{Saggin01},
a modulation quality measurement of microwave access systems presented in \citet{Angrisani10}, 
a real-time rotational speed estimation method using correlation introduced in \citet{Wang14},
a biology-inspired electronic nose developed in \citet{Jing16},
and impedance measurements for material damage estimation method using cross-correlation introduced in \citet{deCastro19}.
The data-driven step input estimation method proposed by \citet{Markovsky15cep} is added to this list, highlighting the fact that it can be implemented for the measurement of different physical magnitudes because it is model-independent.

% Uncertainty assessment 
In metrology, a measurement is an estimation, represented as a random variable.
The measurement noise always exists, and therefore, the input estimation is commonly expressed with its two first statistical moments as it is described in \citet{Ferrero06}.
The guidelines listed in \citet{GUM08}, and accepted by the metrology community, recommend the standardization of the uncertainty assessment.
The typical measurement uncertainty analysis \color{blue} is \color{black} reviewed in \citet{daSilva12}.
The Monte Carlo method is an uncertainty evaluation tool according to \citet{Cox06},
that supports the use of simulation techniques for quantifying measurement uncertainties as suggested by \citet{Esward16}.
One example of the Monte Carlo method application is a dynamic measurement uncertainty evaluation of clinical thermometers described in \citet{Ogorevc16}.

Nevertheless, researchers like \citet{Esward09}, and \citet{Hessling10} still pinpoint the need to study more the uncertainty assessment methods.
\citet{Diniz17} recommends to consider the uncertainties of all the measurement chain components, and to avoid the direct uncertainty propagation from the calibration towards the to-be-measured quantity.  
Methods for evaluating the uncertainty associated with the output of compensation filters have been investigated for
a discrete-time infinite-response filter in \citet{Link09},
a discrete-time finite-response filter in \citet{Elster07, Elster08}, and
the Kalman filter in \citet{Eichstadt16b}.
All these works propagate the uncertainty through the filter, but it is also necessary to upward the propagation up to the sensor model to include all systematic error contributions, as it is explained by \citet{Hessling11}.

@@

This thesis describes research work initiated after the data-driven step input estimation method proposed by \citet{Markovsky15cep}.
This method aims for metrology applications but it lacked an uncertainty assessment, and, therefore, its appropriateness was questionable.
The first research work conducted was a statistical analysis to elucidate the bias and the covariance of the estimate that the data-driven step input estimation method provides.
After obtaining these statistical moments, the estimation uncertainty was assessed, and thus, the effectiveness of the method was appraised.
The second research work was a series of experiments conducted to validate the data-driven step input estimation method in real-life applications.
The experiments were realizations of the step input excitation using temperature and mass sensors.
The step input estimation method showed robustness when the measurement noise is not Gaussian and white, as it was assumed in the theoretical analyses.
The third research work conducted was the estimation of affine input parameters following the data-driven approach for real-time applications. 
The data-driven method uses exponential weighing in the recursive least-squares solution of a structured errors-in-variables problem, to give preference to recent samples over the older samples.
The methodology and performance of the data-driven affine input estimation method were compared to those of a maximum-likelihood method.

@@

The data-driven step input estimation method formulates a structured errors-in-variables (EIV) problem.
In linear estimation EIV problems, the measurement noise perturbs the regression matrix and the regressor, as it is detailed in \citet{VanHuffel91Book} and \citet{Markovsky07overview}.
The regression matrix, in structured EIV problems, has a structure that depends on the problem formulation.
Hankel and Toeplitz matrices appear in several application problems such as those described by \citet{Markovsky15cep} in metrology, by \citet{Soderstrom07} in system identification, by \citet{Feiz17} in image restoration, by \citet{Cai16} in nuclear magnetic resonance spectroscopy, by \citet{Pan18} in direction-of-arrival estimation, and by \citet{Jia18} in time-of-arrival estimation.

The data-driven step input estimation method introduced by \citet{Markovsky15cep} directly estimates the input from the sensor transient response. 
The perturbations in the formulated EIV problem come from the sensor output, since it is the only observed signal, and from this signal both the regression matrix and the regressor are built.
The structure in the regression matrix is block-Hankel.
The data-driven direct estimation methodology  estimates the input faster than the classical two-stage approach described in \citet{Azam15} and \citet{Niedzwiecki16a}, which first identifies a sensor model and later estimates the input using the sensor model.

Instead of using a total least-squares (TLS) solution of the structured EIV problem, the step input estimation method uses the recursive least-squares solution (RLS). 
In the review conducted by \citet{Markovsky07overview} it is shown that the TLS solution of unstructured EIV problems is consistent when the perturbations have zero mean with a given positive definite covariance, and
the TLS solution is equivalent to the maximum likelihood (ML) solution when the disturbances of the EIV problems are i.i.d. normally distributed . 
However, the studies of structured EIV problems presented in \citet{VanHuffel07TLSeditorial} demonstrate that the TLS estimator does not give general results since each structure requires a specific treatment. 
Moreover, \citet{Rhode14recursive} found that the ML estimator leads to non-convex optimization when applied to structured EIV problems, and the global optimum is not guaranteed.
Also, the computational complexity of TLS and ML inhibits their real-time implementation.
Thus, the least-squares (LS) estimator is a suboptimal but simple solution to structured EIV problems that admits a recursive form (RLS) suitable for real-time implementation.

The previously published works that propose LS estimators to solve structured EIV problems do not study the required statistical moments to know the uncertainty of the data-driven step input estimation method.
In the literature we find 
the design of a fast algorithm for matrices with small displacement rank in \citet{Mastronardi07fast}, 
the study of the LS estimator consistency in \citet{Palanthandalam10parameter},
the determination of the bias, and the mean squared error of the parameter estimates in the identification of AR models in \citet{Kiviet12high}, \citet{Kiviet14improved}, and
a discussion of the causes of bias and inconsistency in homogeneous estimators in \citet{Yeredor04homogeneous}.
The literature does not address the uncertainty of LS estimators for structured EIV problems.

The uncertainty of an estimation method is expressed in \citet{Pintelon12Book} using the bias and covariance of the estimate.
To know the uncertainty of the data-driven step input estimation method, we quantified the bias and covariance of the LS solution of EIV problems, for unstructured and structured cases. 
The bias and covariance quantification extend the perturbation analysis that was investigated in \citet{Stewart90SPT} and in \citet{Vaccaro94}.
The conducted analysis provides insight into the impact that the structure and the correlation have on the LS estimation uncertainty.
The study presented in this thesis illustrates a methodology to conduct statistical analysis for any structured EIV problem.

We derived expressions that quantify the bias and covariance by obtaining the mathematical expectation of the LS estimate approximated by the second-order Taylor series expansion.
Using Monte Carlo simulations, we validated the accuracy of the bias and covariance expressions.
These expressions estimate the bias and covariance that the data-driven step input estimation method will exhibit for a given sample size and perturbation level.
We compared in \citet{QuintanaCSDA} the mean squared error of the LS estimate to the minimum variance specified by the Cram\'er-Rao lower bound of the structured EIV problem, to determine the conditions under which the data-driven step input estimation method is appropriate for practical applications.

@@

The validation of the data-driven step input estimation method in a practical application was necessary to demonstrate the usefulness of the method.
The method performance was studied in \citet{Markovsky15cep} using simulations and temperature experiments on a digital signal processor (DSP) of low cost.
The method estimates the unknown level of step inputs by processing the sensor step response, and 
avoiding the sensor modeling stage.
The formulation of the estimation method is a correlated errors-in-variables (EIV) problem with block-Hankel structure.

Other methods for input estimation mainly compensate the sensor transient response, for example, by 
recursive estimation of the compensator parameters in \citet{Shu93}, 
finite impulse response (FIR) filtering in \citet{Elster07}, \citet{Niedzwiecki16b} and 
infinite impulse response (IIR) filtering in \citet{Pintelon90}, \citet{Elster08}.
The uncertainty propagation for these model-based compensators is based on the transfer function or state-space representations of the LTI sensor and filter systems in \citet{Link09} and \citet{Hale09}.
Another way to assess the measurement uncertainty is by observing the results of multiple practical measurements as it is described in \citet{Pietrzak14} for mass and in \citet{Ogorevc16} for temperature sensors.

To validate the data-driven step input estimation method we built a weighing system setup based on a load cell sensor.
The weighing setup was constructed to ensure repeatability and reproducibility, along with different experimental realizations.
Load cell sensors are versatile devices that are applied for
heart and breathing physiological signal monitoring in \citet{Lee16},
clinical analysis of sleep quality in \citet{Zahradka18},
automobile safety studies in \citet{Ballo16},
wind turbine design in \citet{Rossander15}, 
civil engineering structure studies in \citet{Olmi16}, and 
sport bicycle design in \citet{Casas16}, to name a few.

\begin{comment}
% Load cell sensors 
Weighing has been basic for the development of scientific and trade activites. 
The load cell is now a standard transducer for weight determination and also for the improvement of measurement techniques, such as the geometric approach to processing of load cell responses \citet{Kesilmis16}, the design of new conveyor machinery \citet{Yamani18}, and electronic truck scales \citet{Guo18}.

In safety studies, a six axis load cell is devised to quantify accelerations and impact forces exerted on a dummy \citet{Ballo16}.
In alternative energy developments the load cells are useful to measure the forces on the arms of a vertical axis wind turbine \citet{Rossander15}.
An academic study of the load that a structure withstands is conducted with strain gage load cells that confirms the numerical results and facilitates the design of complex shaped structures \citet{Olmi16}.
In sports, the performance of new instrumented crank mechanisms is fostered by the utilization of load cells in the characterization, analysis and validation design stages \citet{Casas16}.

The versatility of the load cells permits the physiological signal monitoring of the the heart and breathing rates \citet{Lee16}, clinical analysis of sleep quality \citet{Zahradka18} and the classification of the movement intensity of people while they are sleeping \citet{Alaziz17}. 
All of these experimental studies have load cells installed on bed setups.
